---
title: "simulationvisualization"
author: "tara"
date: 8/2/22 - Still being updates. 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr)
library(tidyverse)
library(purrr)
library(ggplot2)
library(Metrics)
library(corrplot)
library(kableExtra)
library(metR)
library(cowplot)
library(grid)
library(gridExtra)
library(ggrepel)
library(ggbreak)
```

This creates the hundred data comp visualizations. 

**import the data**
```{r}
phi_Sim1 <-read.csv("Sim1Phis_100data.csv", header=T)%>%mutate(sim="I")
phi_Sim2 <-read.csv("Sim2Phis_100data.csv", header=T)%>%mutate(sim="II")
phi_Sim3 <-read.csv("Sim3Phis_100data.csv", header=T)%>%mutate(sim="III")
fits1<-read.csv("Sim1fitstats_100data.csv", header=T)%>%mutate(sim="I")
fits2<-read.csv("Sim2fitstats_100data.csv", header=T)%>%mutate(sim="II")
fits3 <-read.csv("Sim3fitstats_100data.csv", header=T)%>%mutate(sim="III")
points30 <-read.csv("thirtypoints.csv", header=T)
#points15 <-read.csv("fifteenpoints.csv", header=T)
```

**thirty points comparison figure**  
requested by Tanya  
debated whether to go with R-squared or RMSE and ultimately settled on R-squared because it can be compared across models (unitless)
https://www.statology.org/rmse-vs-r-squared/ 

```{r}
#points30 <-points15 #if you want to look at 15 points really quickly. 

apocolr =c("Total Abundance"="#b2df8a","Hierarchical"="#1f78b4") 

points30$approach <-fct_recode(points30$approach, "Total Abundance"="fit_NT_stats","Hierarchical"="fitsumStats")
#if you want to change to hierarchical all instead of hierarchical summed
#points30$approach <-fct_recode(points30$approach, "Total Abundance"="fit_NT_stats","Hierarchical"="fit_all_Stats")


points30%>%
  filter(rowname=="R2")%>%
  #filter(approach !="fitsumStats")%>%
  ggplot(aes(approach, value, fill=approach))+
  geom_point(aes(y = value, color =approach), 
               position = position_jitterdodge(), 
             size = 1, alpha = 0.2) +
   geom_boxplot(outlier.shape = NA, alpha = 0.8,
                position=position_dodge2(padding=0.5)) + 
 #scale_fill_manual(values = apocolr)+
 # scale_color_manual(values = apocolr)+
 ylab("Out-of-sample r-squared")+xlab("Approach")+
 facet_wrap(~sim, scales="free",ncol=3)+
 guides(fill=guide_legend(title="Analysis"), color=guide_legend(title="Analysis"))+
 theme(
  axis.text.x = element_text(size=12),
  axis.text.y = element_text(size=12),
  axis.title.x=element_text(size=12,face="bold",margin=margin(t=8,unit="pt")),
  axis.title.y=element_text(size=12,face="bold",margin=margin(r=8,unit="pt")),
  strip.text = element_text(size=12),
  legend.key = element_rect(fill = "transparent"),
  legend.text= element_text(size=12),
  legend.title=element_text(size=12, face="bold"),
  legend.position = "bottom",
  strip.background =element_rect(fill="white"),
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "transparent",colour = "dark grey"),
  plot.background = element_rect(fill = "white",colour = "white", size=0.5),
  panel.border = element_rect(colour="black",fill=NA, size=0.3))

#ggsave("thirtypointsfig.png",path="/Users/tdolan/documents/postdoc/age structure/agestructfigs")
#dev.off()
```
If we add them all up (fitsumStats) it's not better than total abundance. 


**Thirty points figure plotted separately and reassembled in InDesign**  
with ggbreak for axis breaks  
https://cran.r-project.org/web/packages/ggbreak/vignettes/ggbreak.html#feature-12-compatible-with-patchwork 
```{r}
# there is probably a way to do scale-break with facets but I am just going to do it manually for now. 
apocolr =c("TA"="#b2df8a","Hier"="#1f78b4") 

points30$approach <-fct_recode(points30$approach, "TA"="Total Abundance","Hier"="Hierarchical")

# 
# 
### I
points30%>%
  filter(rowname=="R2")%>%
  #filter(approach !="fitsumStats")%>%
  filter(approach !="fit_all_stats")%>%
  filter(sim=="I")%>%
  ggplot(aes(approach, value, fill=approach))+
  geom_point(aes(y = value, color =approach), 
               position = position_jitterdodge(), 
             size = 1, alpha = 0.8) +
   geom_boxplot(width=0.3, outlier.shape = NA, alpha = 0.8,
                position=position_dodge2(padding=0.5)) + 
 scale_fill_manual(values = apocolr)+
  scale_color_manual(values = apocolr)+
 ylab("")+xlab("")+
 #facet_wrap(~sim, scales="free",ncol=3)+
 #guides(fill=guide_legend(title="Analysis"), color=guide_legend(title="Analysis"))+
 theme(
  axis.text.x = element_text(size=16),
  axis.text.y = element_text(size=16),
  axis.title.x=element_text(size=16,face="bold",margin=margin(t=8,unit="pt")),
  axis.title.y=element_text(size=16,face="bold",margin=margin(r=8,unit="pt")),
  strip.text = element_text(size=16),
  legend.key = element_rect(fill = "transparent"),
  legend.text= element_text(size=16),
  legend.title=element_text(size=16, face="bold"),
  legend.position = "none",
  strip.background =element_rect(fill="white"),
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "transparent",colour = "dark grey"),
  plot.background = element_rect(fill = "white",colour = "white", size=0.5),
  panel.border = element_rect(colour="black",fill=NA, size=0.3))
#ggsave("thirtypointsfigI.png",height=6, width=4,dpi=300, path="/Users/tdolan/documents/postdoc/age structure/agestructfigs")

### II
points30%>%
  filter(rowname=="R2")%>%
  #filter(approach !="fitsumStats")%>%
  filter(approach !="fit_all_stats")%>%
  filter(sim=="II")%>%
  ggplot(aes(approach, value, fill=approach))+
  geom_point(aes(y = value, color =approach), 
               position = position_jitterdodge(), 
             size = 1, alpha = 0.8) +
   geom_boxplot(width=0.3, outlier.shape = NA, alpha = 0.8,
                position=position_dodge2(padding=0.5)) + 
 scale_fill_manual(values = apocolr)+
  scale_color_manual(values = apocolr)+
  #scale_y_break(c(0.5,0.7), scales=5)+
  scale_y_break(c(0.8,0.925), scales=2)+
 ylab("")+xlab("")+
 #facet_wrap(~sim, scales="free",ncol=3)+
 #guides(fill=guide_legend(title="Analysis"), color=guide_legend(title="Analysis"))+
 theme(
  axis.text.x = element_text(size=16),
  axis.text.y = element_text(size=16),
  axis.title.x=element_text(size=16,face="bold",margin=margin(t=8,unit="pt")),
  axis.title.y=element_text(size=16,face="bold",margin=margin(r=8,unit="pt")),
  strip.text = element_text(size=16),
  legend.key = element_rect(fill = "transparent"),
  legend.text= element_text(size=16),
  legend.title=element_text(size=16, face="bold"),
  legend.position = "none",
  strip.background =element_rect(fill="white"),
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "transparent",colour = "dark grey"),
  plot.background = element_rect(fill = "white",colour = "white", size=0.5),
  panel.border = element_rect(colour="black",fill=NA, size=0.3))
#ggsave("thirtypointsfigII.png",height=6, width=4,dpi=300, path="/Users/tdolan/documents/postdoc/age structure/agestructfigs")

### III
points30%>%
  filter(rowname=="R2")%>%
  filter(approach !="fit_all_stats")%>%
  #filter(approach !="fitsumStats")%>%
  filter(sim=="III")%>%
  ggplot(aes(approach, value, fill=approach))+
  geom_point(aes(y = value, color =approach), 
               position = position_jitterdodge(), 
             size = 1, alpha = 0.8) +
   geom_boxplot(width=0.3, outlier.shape = NA, alpha = 0.8,
                position=position_dodge2(padding=0.5)) + 
 scale_fill_manual(values = apocolr)+
  scale_color_manual(values = apocolr)+
  #scale_y_break(c(0.94,0.96), scales=2)+
  scale_y_break(c(0.36,0.92), scales=5)+
 ylab("")+xlab("")+
 #facet_wrap(~sim, scales="free",ncol=3)+
 #guides(fill=guide_legend(title="Analysis"), color=guide_legend(title="Analysis"))+
 theme(
  axis.text.x = element_text(size=16),
  axis.text.y = element_text(size=16),
  axis.title.x=element_text(size=16,face="bold",margin=margin(t=8,unit="pt")),
  axis.title.y=element_text(size=16,face="bold",margin=margin(r=8,unit="pt")),
  strip.text = element_text(size=16),
  legend.key = element_rect(fill = "transparent"),
  legend.text= element_text(size=16),
  legend.title=element_text(size=16, face="bold"),
  legend.position = "none",
  strip.background =element_rect(fill="white"),
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "transparent",colour = "dark grey"),
  plot.background = element_rect(fill = "white",colour = "white", size=0.5),
  panel.border = element_rect(colour="black",fill=NA, size=0.3))
#ggsave("thirtypointsfigIII.png",height=6, width=4,dpi=300, path="/Users/tdolan/documents/postdoc/age structure/agestructfigs")
```
We notice that when you use the fitsumStats (where the obs and pred are added up from each model) total abundance is BETTER than not. Whereas if you use the overall R2 (direct from package) Total abundance is slightly less good or basically equivalent. 

**Tileplot of ARD values for hyperparameters**
```{r}
#create giant dataframe from all simulations
phis <-bind_rows(phi_Sim1, phi_Sim2, phi_Sim3)%>%dplyr::select(-X)%>%rename(models=rowname)
phis <-phis[,c(1,2,3,7,8,9,4,5,6,10,12,13,14,15,11)]
pivphi <-pivot_longer(phis, 2:6, names_to = "name")%>%as.data.frame()

#create pivot version where it's the mean only. 
pivphimean <-pivphi %>% filter(approach != "TAindex")%>%group_by(name,model,sim)%>% summarize(phimean=mean(value),phisd=sd(value),.groups="drop")
pivphimean[is.na(pivphimean)] <-0
#pivphimean <-mutate(pivphimean, tslength=ifelse(rowname=="N_total",100,substr(rowname,1,2)))%>%mutate(tslength=ifelse(tslength=="5y",5,tslength))%>%mutate(maxE=round(sqrt(as.numeric(tslength))))

#discretize the color scale
cols <- c("(0,0.2]"="#f1eef6", "(0.2,0.4]" = "#d0d1e6", "(0.4,0.6]" = "#a6bddb", "(0.6,0.8]" = "#74a9cf", "(0.8,1.0]"  = "#2b8cbe", "(1.0,1.2]"="#045a8d", "(1.2,1.4]" ="#034e7b", "(1.4,1.6]" ="#011f4b")
pivphimean<-mutate(pivphimean, phimean_cut = cut(phimean, breaks=c(0,0.2,0.4,0.6,0.8,1.0,1.2,1.4,1.6)))

# need to fix the legend on this plot. Also maybe the low value color shouldn't be white because it's hard to see. 
ggplot(pivphimean, aes(name,model))+
 geom_tile(aes(fill=phimean_cut), show.legend = TRUE)+
 #scale_fill_gradient(low="white",high="blue")+
 scale_fill_manual(values=cols, na.value = "white")+
 xlab("")+ylab("")+
        #+coord_flip()+
 guides(fill=guide_legend(title="mean value of phi"))+
 facet_wrap(~sim)+
 theme_bw()+
 theme(panel.background = element_rect(fill = "transparent", colour = "black"),
       #axis.text.x = element_text(angle = 90),
       axis.text.x=element_text(size=12),
       axis.text.y=element_text(size=12),
       strip.background =element_rect(fill="white"),
       panel.grid.major = element_blank(), 
       panel.grid.minor = element_blank())
#ggsave("phiplot.png", path="/Users/tdolan/documents/postdoc/age structure/agestructfigs")
#ggsave("phiplot.png")
#dev.off()
```
**The different approaches**  
hier = normal hierarchical GP fit. The age classes and time series length are what you would expect based on the "models" column.  
single_age = one age at a time, for as many age classes as expected in the model, 100 years of data. 
sumhier = another way of aggregating the hierarchical results. Instead of using the overall prediction aggregated within the r package, take the prediction from the hiearchical model for each age class, first get the observed and predicted. unlog them, add them, then log them again. I am not sure why we did this. 
TAindex = an index of total abundance comprised of the length and number of age classes listed in the models. 

**pivot**
```{r}
newfits <-bind_rows(fits1,fits2,fits3)%>%dplyr::select(-X)

newfits <-mutate(newfits, tslength=ifelse(!is.na(tslength),tslength,substr(model,nchar(model)-1,nchar(model))),tslength)%>%
  mutate(tslength=ifelse(tslength=="05",5,tslength))


#pivot
nfitspiv <-pivot_longer(newfits, 2:9, names_to = "stat")%>%as.data.frame()%>%mutate(model=as.character(model), sim=as.factor(sim))%>%
  mutate(maxE=round(sqrt(as.numeric(tslength))))

#The Hier version
fig2filt <-filter(nfitspiv,stat=="OOS_R2", approach %in% c("TAindex","hier"), model !="preyNT")%>%mutate(tslength=as.factor(tslength))
fig2filt$tslength <-fct_relevel(fig2filt$tslength,c("5","10","20","25"))

#the SumHier version
fig4filt <-filter(nfitspiv,stat=="OOS_R2", 
                  approach %in% c("TAindex","sumhier"), 
                  model %in%c("prey205","prey1010","prey520","prey425"))%>% 
                 # age_class %in% c("4","5","10","20","all"))%>%
  mutate(tslength=as.factor(tslength))
fig4filt$tslength <-fct_relevel(fig4filt$tslength,c("5","10","20","25"))

```

**Boxplot1: Figure total abundance single index to overall estimates from hierarchical models**
under "approach", "sumhier" is the sum total abundance of the hierarchcial models. 
LATER we found that we do not want to compare the summed total abundance estimates to total abundance, but we want to compare overall r-squared to total abundance. So is that "hier"


**Boxplot1: but one at a time**
Boxplot of age structured model next to Ntotal model. 
```{r}
mods1 <-c("prey1010","prey205","prey425","prey520")
tslength<-c(5,10,20,25)

apocolr =c("TAindex"="#b2df8a","sumhier"="#1f78b4")


plotlist <-list()
jlist <-list()
simlist <-c("I","II","III")
for (j in 1:length(simlist)){
for (i in 1:length(tslength)){
    #fits <-filter(fig2filt, tslength==tslength[i])%>%
       fits <-filter(fig4filt, tslength==tslength[i])%>%
    filter(sim==simlist[j])%>%filter(stat %in% c("OOS_R2"))
    #filter(sim=="III")
    means <-fits %>% group_by(approach)%>%summarize(avr2=mean(value))
  ts <-tslength[i]
 p<- fits%>%arrange(tslength)%>%
  ggplot(aes(approach, value,fill=approach))+
  #geom_boxplot(aes(fill=model),alpha=0.7,notch=TRUE, notchwidth = 0.8,width=0.5)+ 
    #geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
    geom_point(aes(y = value, color = approach), 
              #position = position_jitter(width = 0.15), size = 1, alpha = 0.2) +
   position = position_jitterdodge(dodge.width=2), size = 1, alpha = 0.5) +
    geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.8,
                 position=position_dodge2()) +
   geom_line(data = means,aes(x=approach, y=avr2))+
 scale_fill_manual(values = apocolr)+
scale_color_manual(values = apocolr)+
   ylim(min(fits$value),max(fits$value))+
 ylab("")+xlab("")+#ggtitle(ts)+
 #guides(fill=guide_legend(title="model"))+
 guides(color=FALSE, fill=FALSE)+
   theme_bw()+
 theme(
  #axis.text.x = element_text(size=14),
  axis.text.x = element_blank(),
  axis.text.y = element_text(size=14),
  strip.text = element_text(size=14),
  strip.background =element_rect(fill="transparent"),
  legend.position = "none",
  plot.margin=unit(c(1,1,1,1),"pt"), #top right bottom left
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "transparent",colour = "dark grey"),
  plot.background = element_rect(fill = NA,colour = "transparent", size=0.5),
  panel.border = element_rect(colour="transparent",fill=NA, size=0.3))
 plotlist[[i]] <-p
}
 jlist[[j]]<-plotlist
}

plotlist1 <-plot_grid(jlist[[1]][[1]], jlist[[1]][[2]], jlist[[1]][[3]],jlist[[1]][[4]], 
            ncol = 4 #,labels=c("a.","b.","c.","d.")
                      )
plotlist2 <-plot_grid(jlist[[2]][[1]], jlist[[2]][[2]], jlist[[2]][[3]],jlist[[2]][[4]], 
                      ncol = 4 #,labels=c("e.","f.","g.","h.")
                      )
plotlist3 <-plot_grid(jlist[[3]][[1]], jlist[[3]][[2]], jlist[[3]][[3]],jlist[[3]][[4]], 
                      ncol = 4 #,labels=c("i.","j.","k.","l.")
                      )

y.grob <- textGrob("Out-of-sample r-squared", 
                   gp=gpar(fontface="bold", col="black", fontsize=12), rot=90)

x.grob <- textGrob("", 
                   gp=gpar(fontface="bold", col="black", fontsize=12))


mainplot <-plot_grid(plotlist1,plotlist2,plotlist3, nrow = 3)
grid.arrange(arrangeGrob(mainplot, left = y.grob, top = x.grob,padding = unit(20, "line")))


#ggsave("cplotattempt4sumhier.png", height=12, width=10, path="/Users/tdolan/documents/postdoc/age structure/agestructfigs")
#dev.off()
```

**Boxplot1: Tanya's version***
My thought is to have 3 panels stacked vertically, one for each simulation, each with its own vertical axis. Horizontal axis (shared) is time series length (years). At each horizontal axis tick there are 2 boxplots (or points and error bars) of different colors side by side (you may have to map color rather than fill if some end up being small). One color is 'with age structure', the other is 'total abundance' indicated in legend. Caption will explain that the models with age structure use the number of age classes such that number of years*number of age classes = 100 data points (see main text), or something like that. Explain that the variation shown is among 100 replicate simulations.
```{r}
apocolr =c("TAindex"="#b2df8a","hier"="#1f78b4")

fig2filt%>%
  ggplot(aes(tslength, value, fill=approach))+
  geom_point(aes(y = value, color =approach), 
               position = position_jitterdodge(), 
             size = 1, alpha = 0.2) +
   geom_boxplot(outlier.shape = NA, alpha = 0.8,
                position=position_dodge2(padding=0.5)) + 
   #geom_violin()+
 scale_fill_manual(values = apocolr)+
  scale_color_manual(values = apocolr)+
 ylab("Out-of-sample r-squared")+xlab("Time series length (years)")+
 facet_wrap(~sim, scales="free",ncol=3)+
 guides(fill=guide_legend(title="Analysis"), color=guide_legend(title="Analysis"))+
 theme(
  axis.text.x = element_text(size=12),
  axis.text.y = element_text(size=12),
  axis.title.x=element_text(size=12,face="bold",margin=margin(t=8,unit="pt")),
  axis.title.y=element_text(size=12,face="bold",margin=margin(r=8,unit="pt")),
  strip.text = element_text(size=12),
  legend.key = element_rect(fill = "transparent"),
  legend.text= element_text(size=12),
  legend.title=element_text(size=12, face="bold"),
  legend.position = "bottom",
  strip.background =element_rect(fill="white"),
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "transparent",colour = "dark grey"),
  plot.background = element_rect(fill = "white",colour = "white", size=0.5),
  panel.border = element_rect(colour="black",fill=NA, size=0.3))

ggsave("Boxplot_Hier_TA.png",path="/Users/tdolan/documents/postdoc/age structure/agestructfigs")
dev.off()


```



**Boxplot1: Same as above but use sum hier instead of hier***
```{r}
apocolr =c("TAindex"="#b2df8a","sumhier"="#1f78b4")

fig4filt <-filter(nfitspiv,stat=="OOS_R2", 
                  approach %in% c("TAindex","sumhier"), 
                  model %in%c("prey205","prey1010","prey520","prey425"))%>% 
                 # age_class %in% c("4","5","10","20","all"))%>%
  mutate(tslength=as.factor(tslength))
fig4filt$tslength <-fct_relevel(fig4filt$tslength,c("5","10","20","25"))


fig4filt%>%
  ggplot(aes(tslength, value, fill=approach))+
  geom_point(aes(y = value, color =approach), 
               position = position_jitterdodge(), 
             size = 1, alpha = 0.2) +
   geom_boxplot(outlier.shape = NA, alpha = 0.8,
                position=position_dodge2(padding=0.5)) + 
   #geom_violin()+
 scale_fill_manual(values = apocolr)+
  scale_color_manual(values = apocolr)+
 ylab("Out-of-sample r-squared")+xlab("Time series length (years)")+
 facet_wrap(~sim, scales="free",ncol=3)+
 guides(fill=guide_legend(title="Analysis"), color=guide_legend(title="Analysis"))+
 theme(
  axis.text.x = element_text(size=12),
  axis.text.y = element_text(size=12),
  axis.title.x=element_text(size=12,face="bold",margin=margin(t=8,unit="pt")),
  axis.title.y=element_text(size=12,face="bold",margin=margin(r=8,unit="pt")),
  strip.text = element_text(size=12),
  legend.key = element_rect(fill = "transparent"),
  legend.text= element_text(size=12),
  legend.title=element_text(size=12, face="bold"),
  legend.position = "bottom",
  strip.background =element_rect(fill="white"),
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "transparent",colour = "dark grey"),
  plot.background = element_rect(fill = "white",colour = "white", size=0.5),
  panel.border = element_rect(colour="black",fill=NA, size=0.3))

ggsave("Boxplot_sumhier_TA.png",path="/Users/tdolan/documents/postdoc/age structure/agestructfigs")
dev.off()


```

**Boxplot 2: 100 data points test**
hierarchical models of 100 data points. 
```{r}
fig3filt <-filter(nfitspiv,stat=="OOS_R2", approach %in% c("TAindex","hier"), 
                  model %in% c("preyNT","prey205","prey1010","prey520","prey425"),
                  age_class =="all")%>% 
  mutate(tslength=as.factor(tslength))
fig3filt$tslength <-fct_relevel(fig3filt$tslength,c("5","10","20","25","100"))
fig3filt$model <-fct_relevel(fig3filt$model,c("prey205","prey1010","prey520","prey425","preyNT"))

fig3filt%>%
  ggplot(aes(tslength, value))+
  geom_point(aes(y = value), color="#1f78b4",
               position = position_jitter(width=0.2), size = 1, alpha = 0.4) +
    geom_boxplot(width = 0.3, outlier.shape = NA, alpha = 0.8,fill="#1f78b4") +
    #geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
 ylab("R-squared")+
 facet_wrap(~sim, scales="free_y")+
 guides(fill=guide_legend(title="analysis"))+
 theme(
  axis.text.x = element_text(size=12),
  axis.text.y = element_text(size=12),
  axis.title.x=element_text(size=12,face="bold",margin=margin(t=8,unit="pt")),
  axis.title.y=element_text(size=12,face="bold",margin=margin(r=8,unit="pt")),
  strip.text = element_text(size=12),
  legend.key = element_rect(fill = "transparent"),
  legend.text= element_text(size=12),
  legend.title=element_text(size=12, face="bold"),
  legend.position = "bottom",
  strip.background =element_rect(fill="white"),
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "transparent",colour = "dark grey"),
  plot.background = element_rect(fill = "white",colour = "white", size=0.5),
  panel.border = element_rect(colour="black",fill=NA, size=0.3))


ggsave("Boxplot_hier.png", path="/Users/tdolan/documents/postdoc/age structure/agestructfigs")
dev.off()
```


**summary dynamic correlation**
```{r}
sumfits <-nfitspiv %>% 
  filter(stat=="rho" & approach=="hier")%>%
  filter(tslength != "100")%>%
  group_by(sim, model)%>% summarize('mean dynamic rho'=mean(value), "std. dev. dynamic rho"=sd(value), .groups="drop")%>%mutate(across(c(3,4), round,3))



sumfits%>%arrange(model,sim)%>% kbl()%>%kable_classic(full_width = F, html_font = "Cambria")

```


**measure dynamic correlation in each simulation**
```{r}

nfitspiv %>%
 filter(stat == "rho")%>%
  filter(model != "preyNT")%>%
 ggplot(aes(models,value, fill=models))+
 geom_boxplot(notch=T, notchwidth = 0.8)+
 xlab("model")+ylab("mean dynamic rho")+
 #scale_fill_manual(values=simcols3)+
 facet_wrap(~sim)+
 theme(
  axis.text.x = element_text(size=9, angle=90),
  axis.text.y = element_text(size=7),
  strip.background =element_rect(fill="transparent"),
  strip.text = element_text(size=9),
  legend.position = "none",
  panel.spacing = unit(1.5, "mm"),
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "transparent",colour = "dark grey"),
  plot.background = element_rect(fill = NA,colour = "black", size=0.5),
  panel.border = element_rect(colour="black",fill=NA, size=0.3)
 )

#ggsave("meandynamicrho.png", path="/Users/tdolan/documents/postdoc/age structure/agestructfigs")
#dev.off()
```









